---
title: "1. Tiered quarantine: outbreak probability tables"
subtitle: "Measles in Grantz schools, real SMART per-capita contact matrices, R0 = 12"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    df-print: kable
    embed-resources: true
  gfm:
    echo: false        # GitHub version shows results only; code is in the .qmd
    toc: true
    toc-depth: 2
execute:
  warning: false
  message: false
---

Tiered-quarantine experiment using the three real per-capita SMART contact
matrices (symmetrised for reciprocity, kept at real magnitude). R0 = 12 is
reached through a per-school transmission rate, not by scaling the matrix.

For each school we report **P(outbreak e 10 / 20 / 50 cases)** across
vaccination coverage and quarantine strategies, written as
**high / medium / low-risk contact quarantine days**.

The diagnostic single-tier scenarios in the first table do not satisfy
high e medium e low; that constraint is applied to every other analysis.

**Difference from the reference.** Each strategy is compared with a reference
scenario in the same school and at the same vaccination coverage: *No
quarantine* for the no-vaccination table, and *21/21/21* for the tiered
tables. The "Difference" columns give the strategy's outbreak probability
minus the reference's, in **percentage points**, with a 95% confidence
interval (Newcombe hybrid score method for a difference of two proportions).

- A **positive** difference means more outbreaks than the reference; a
  **negative** one means fewer.
- If the interval **includes 0**, the strategy cannot be told apart from the
  reference with this number of simulations.

Every school × scenario is simulated with its own random seed, so the
scenarios are independent samples, as the interval assumes. The interval
reflects simulation (Monte Carlo) uncertainty only, not uncertainty in the
model's assumptions.

::: {.callout-note}
Model code lives in `measles_model.R`. Simulations are cached in
`results/`; delete the CSV to re-run. For a quick test render with
`N_SIMS=200 quarto render 01_tiered_quarantine_tables.qmd`.
:::

::: {.content-hidden when-format="markdown"}

## Setup

```{r setup}
source("measles_model.R")

THRESHOLDS   <- c(10, 20, 50)                       # P(size >= K)
VACCS_TIERED <- c(0.50, 0.60, 0.70, 0.80, 0.90, 0.95)
REFERENCE    <- c(novac = "No quarantine", tiered = "21/21/21")
```

:::

## Transmission rate check

Each school gets its own transmission rate so that R0 = 12. The table checks
this against `epiworldR::compute_reproduction_number()`.

```{r r0-check}
r0_check <- rbindlist(lapply(SCHOOLS, function(s) data.table(
  School              = s,
  Students            = sum(ENROLLMENT[[s]]),
  `Transmission rate` = round(TRANS_RATE[[s]], 4),
  R0 = round(epiworldR::compute_reproduction_number(
         contact_matrix         = SCHOOL_CMAT[[s]],
         transmission_prob      = TRANS_RATE[[s]],
         infectious_period_days = INF_DAYS,
         check_reciprocity      = TRUE)$R, 2)
)))
knitr::kable(r0_check)
```

::: {.content-hidden when-format="markdown"}

## Scenarios

```{r design}
## A. No vaccination: which single tier is worth quarantining? (diagnostic)
design_novac <- strategy_table(c(
  "No quarantine"        = "0/0/0",
  "Only high-risk (21d)" = "21/0/0",
  "Only med-risk (21d)"  = "0/21/0",
  "Only low-risk (21d)"  = "0/0/21"))
design_novac[, vacc := 0]

## B. Tiered strategies (all high >= medium >= low) at each coverage
design_tiered <- cross_vacc(
  strategy_table(c("21/21/21", "21/14/14", "21/7/7", "21/0/0")),
  VACCS_TIERED)

design <- rbind(data.table(set = "novac",  design_novac),
                data.table(set = "tiered", design_tiered))
```

```{r simulate}
## Each school x scenario gets its own seed, so scenarios are independent
## samples (needed for the confidence intervals below)
sims <- run_or_load("tables_indep_seeds",
                    simulate_design(design, independent_seeds = TRUE))
```

```{r tabulate}
## Number of runs with an outbreak (size >= K) for each threshold
counts <- sims[, c(.(n = .N),
                   setNames(lapply(THRESHOLDS, function(k) sum(size >= k)),
                            paste0("x", THRESHOLDS))),
               by = .(set, school, vacc, strategy)]

## Attach the reference scenario's counts (same set, school and coverage)
counts[, ref_name := REFERENCE[set]]
ref <- counts[strategy == ref_name,
              c(list(set = set, school = school, vacc = vacc, n_ref = n),
                setNames(.SD, paste0("xr", THRESHOLDS))),
              .SDcols = paste0("x", THRESHOLDS)]
counts <- ref[counts, on = .(set, school, vacc)]

## Difference of two independent proportions, p_new - p_ref,
## with Newcombe's hybrid score 95% CI (built from two Wilson intervals)
newcombe_diff <- function(x_new, n_new, x_ref, n_ref) {
  p1 <- x_new / n_new;  p2 <- x_ref / n_ref
  w1 <- wilson(x_new, n_new);  w2 <- wilson(x_ref, n_ref)
  d  <- p1 - p2
  list(diff = d,
       lo   = d - sqrt((p1 - w1$lo)^2 + (w2$hi - p2)^2),
       hi   = d + sqrt((w1$hi - p1)^2 + (p2 - w2$lo)^2))
}

fmt_pct  <- function(p) sprintf("%.1f%%", 100 * p)
fmt_diff <- function(d) {
  s <- function(v) sprintf("%+.1f", 100 * v)
  sprintf("%s (%s to %s)", s(d$diff), s(d$lo), s(d$hi))
}

PROB_COLS <- paste0("P(\u2265", THRESHOLDS, ")")
DIFF_COLS <- paste0("Difference P(\u2265", THRESHOLDS, ")")

for (i in seq_along(THRESHOLDS)) {
  x  <- counts[[paste0("x",  THRESHOLDS[i])]]
  xr <- counts[[paste0("xr", THRESHOLDS[i])]]
  counts[, (PROB_COLS[i]) := fmt_pct(x / n)]
  counts[, (DIFF_COLS[i]) := ifelse(strategy == ref_name, "Reference",
                                    fmt_diff(newcombe_diff(x, n, xr, n_ref)))]
}

counts[, strategy := factor(strategy, levels = unique(design$strategy))]
setorder(counts, school, vacc, strategy)

## One results table
results_table <- function(d, first_col = "Strategy") {
  knitr::kable(d[, c("strategy", PROB_COLS, DIFF_COLS), with = FALSE],
               col.names = c(first_col, PROB_COLS, DIFF_COLS),
               align = c("l", rep("r", length(PROB_COLS)), rep("r", length(DIFF_COLS))))
}
```

:::

## Results by school

Differences are in percentage points with 95% confidence intervals.
For example, **+3.1 (+1.2 to +5.0)** means the strategy has 3.1 percentage
points more chance of that outbreak size than the reference.

```{r results, results='asis'}
for (s in SCHOOLS) {
  cat("\n\n### ", s, "\n\n")

  cat("**No vaccination: which single tier to quarantine** (compared with no quarantine)\n\n")
  print(results_table(counts[school == s & set == "novac"], "Scenario"))

  cat("\n\n**Tiered strategies (high/medium/low days)**, compared with 21/21/21 at the same coverage\n\n")
  ## A separate table per coverage level: a label inside a Markdown table
  ## would end the table, so each coverage level gets its own
  for (v in VACCS_TIERED) {
    cat(sprintf("\n\n*%s vaccinated*\n\n", vacc_lab(v)))
    print(results_table(counts[school == s & set == "tiered" & vacc == v]))
  }
}
```

