---
title: "2. Outbreak-size distributions by quarantine strategy"
subtitle: "Measles in Grantz schools, real SMART per-capita contact matrices, R0 = 12"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    df-print: kable
    embed-resources: true
  gfm:
    echo: false        # GitHub version shows results only; code is in the .qmd
    toc: true
    toc-depth: 2
execute:
  warning: false
  message: false
---

How big do outbreaks get under each quarantine strategy? For every
vaccination level and one of 11 strategies (**high / medium / low-risk
quarantine days**), we look at the full distribution of final outbreak sizes
rather than a single probability.

- **Histograms** are drawn as outlines, one coloured line per school, so the
  three schools can be compared in the same panel. There is one panel per
  strategy and one graph per vaccination level.
- **Tables** give the probability of an outbreak of at least 10, 20 and 50
  cases, for each school and strategy.

::: {.callout-note}
Model code lives in `measles_model.R`. Simulations are cached in
`results/`, and PNG copies of every figure are saved to `figures/`.
:::

::: {.content-hidden when-format="markdown"}

## Setup

```{r setup}
source("measles_model.R")

K        <- 20                 # dashed line in the histograms
TABLE_KS <- c(10, 20, 50)      # table: P(outbreak size >= each of these)
VACCS <- c(0.50, 0.60, 0.70, 0.80, 0.85)      # one set of figures each

STRATS <- c("21/21/21", "21/14/7", "21/7/7", "21/7/0",
            "14/14/0",  "14/7/7",  "14/7/0", "14/0/0",
            "7/7/0",    "7/0/0",   "0/0/0")

HIST_BINWIDTH <- 5           # outbreak-size bin width (cases)

## Outline colours, one per school. Any R colour name ("red", "darkgreen")
## or hex code ("#E69F00") works. Default: the same colours as file 3.
HIST_COLORS <- SCHOOL_COLORS
## Examples:
## HIST_COLORS <- c(Elementary = "#D55E00", Middle = "#000000", High = "#56B4E9")
HIST_COLORS <- c(Elementary = "#D55E00", Middle = "#0072B2", High = "#000000")    # 1. vermillion / blue / black (Okabe-Ito)
# HIST_COLORS <- c(Elementary = "#4477AA", Middle = "#EE6677", High = "#228833")  # 2. blue / red / green (Paul Tol)
HIST_LINEWIDTH <- 0.8        # outline thickness
HIST_FACET_Y  <- "free_y"    # "fixed" to compare bar heights across panels
```

## Simulations

Each school &times; vaccination &times; strategy combination is simulated once and reused
by every figure and table below.

```{r simulate}
design <- cross_vacc(strategy_table(STRATS), VACCS)
sims   <- run_or_load("distributions", simulate_design(design))
sims[, strategy := factor(strategy, levels = STRATS)]
```

```{r plot-functions}
## Histogram outlines: for each school, the number of runs in each bin,
## drawn as a step line (the outline a histogram's bars would have).
hist_outline <- function(d) {
  bw   <- HIST_BINWIDTH
  edge <- seq(0, ceiling((max(sims$size) + 1) / bw) * bw, by = bw)
  d[, left := floor(size / bw) * bw]
  counts <- d[, .N, by = .(school, strategy, left)]
  ## every bin for every school x strategy, including empty ones (count 0)
  grid <- CJ(school = unique(d$school), strategy = unique(d$strategy),
             left = head(edge, -1))
  counts <- counts[grid, on = .(school, strategy, left)][is.na(N), N := 0L]
  setorder(counts, school, strategy, left)
  ## step path: start at 0, follow each bin's count, drop back to 0 at the end
  counts[, {
    .(x = c(0, left, max(edge), max(edge)),
      y = c(0, N, last(N), 0))
  }, by = .(school, strategy)]
}

plot_hist <- function(v) {
  out <- hist_outline(copy(sims[vacc == v]))
  ggplot(out, aes(x, y, colour = school)) +
    geom_step(direction = "hv", linewidth = HIST_LINEWIDTH) +
    geom_vline(xintercept = K, linetype = "longdash",
               linewidth = 0.6, colour = "grey45") +
    facet_wrap(~ strategy, ncol = 3, scales = HIST_FACET_Y) +
    scale_colour_manual(values = HIST_COLORS, name = "School") +
    ## same x-range in every panel and every graph
    coord_cartesian(xlim = c(0, max(sims$size) + HIST_BINWIDTH), expand = FALSE) +
    labs(x = "Final outbreak size", y = "Number of runs",
         title = sprintf("Outbreak-size distribution by strategy, %s vaccinated",
                         vacc_lab(v)),
         subtitle = sprintf(paste0("Histogram outlines, bins of %d cases. ",
                                   "Panels = high/medium/low quarantine days. Dashed line = %d cases"),
                            HIST_BINWIDTH, K)) +
    theme_measles(11) +
    theme(legend.position = "top", panel.grid.major.x = element_blank())
}


```

:::

## Histograms, all schools

::: {.panel-tabset}

```{r histograms, results='asis', fig.width=11, fig.height=10}
for (v in VACCS) {
  cat("\n\n## ", vacc_lab(v), " vaccinated\n\n", sep = "")

  p <- plot_hist(v)

  print(save_fig(
    p,
    sprintf("fig_hist_all_schools_%02d.png", round(100 * v)),
    11, 10
  ))

  cat("\n\n")

}
```
