---
title: "3. Which quarantine policy keeps outbreaks small?"
subtitle: "Measles in Grantz schools, real SMART per-capita contact matrices, R0 = 12"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    df-print: kable
    embed-resources: true
    fig-dpi: 150
  gfm:
    echo: false        # GitHub version shows results only; code is in the .qmd
    toc: true
    toc-depth: 2
execute:
  warning: false
  message: false
---

This is the main analysis. Every combination of school, vaccination coverage
and **high / medium / low-risk quarantine days** (with high &ge; medium &ge; low) is
simulated once, and all figures are computed from that single set of runs.

| Figure | Question |
|:--|:--|
| 1  | How risky is each school with minimal quarantine? |
| 2  | What happens to outbreak risk when one factor changes, and is the change clear? |
| 3  | Which quarantine combinations keep risk low? |
| 4  | Given high- and medium-risk days, how many low-risk days are worth it? |
| 5  | How big do outbreaks get, with high-risk-only and with optimal quarantine? |
| 6  | How does risk change with low-risk days, across thresholds? |

Plain-language conclusions, computed from the results, are at the end and are
also written to `results/conclusions.txt`.

::: {.callout-note}
Model code lives in `measles_model.R`. Simulations are cached in
`results/`, and PNG copies of every figure are saved to `figures/`.
:::

::: {.content-hidden when-format="markdown"}

## Setup

```{r setup}
source("measles_model.R")

## Design grid
VACCS <- c(0.60, 0.70, 0.80, 0.85)
HIGHS <- c(7L, 14L, 21L)
MEDS  <- c(0L, 7L, 14L, 21L)
LOWS  <- c(0L, 3L, 5L, 7L, 10L, 14L, 21L)

## Analysis choices
K_MAIN     <- 25L                  # an "outbreak" = MORE THAN K_MAIN cases
KS         <- c(10L, 35L, 50L)     # thresholds compared in Figure 6
TARGETS    <- c(0.10, 0.20, 0.30)  # acceptable outbreak risks (Figure 3)
OPT_SHARE  <- 0.80   # "optimal" = fewest days giving 80% of the possible benefit
MIN_GAIN   <- 0.02   # if quarantine can't cut risk by 2 points, optimal = 0 days
FIG3_LOWS  <- c(0L, 7L, 14L, 21L)  # low-risk days shown in Figure 3
FIG2_TABLE_BENEFIT <- 0.05  # Figure 2 table: only comparisons with a benefit under 5 points
FIG3_COLORS    <- c("#fff5eb", "#fdae6b", "#d94801", "#7f2704")  # Figure 3: low -> high risk
FIG3_MAX       <- NA     # top of Figure 3's colour scale; NA = highest risk shown, 1 = 100%
FIG3_LABELS    <- TRUE   # print the risk (%) in each square
FIG3_HIGHLIGHT <- 0.10   # outline squares with risk at or below this; NA = no outlines
ADD_CAPTION <- FALSE # TRUE adds a methods caption under every saved figure
```

```{r plot-settings}
theme_set(theme_measles(12))

ylab_risk  <- sprintf("Chance of more than %d cases", K_MAIN)
VACC_LABS  <- vacc_lab(VACCS)
MIN_POLICY <- sprintf("%d / 0 / 0", min(HIGHS))

CAPTION <- sprintf(
  "Outbreak = more than %d cases. %s simulations per setting, each starting with one infected student.",
  K_MAIN, format(N_SIMS, big.mark = ","))

## save_fig() plus the optional caption
fig <- function(p, file, width, height) {
  if (ADD_CAPTION)
    p <- p + labs(caption = CAPTION) +
      theme(plot.caption = element_text(colour = "grey35", size = 9, hjust = 0))
  save_fig(p, file, width, height)
}
```

:::

## Simulations

```{r simulate}
design <- policy_grid(VACCS, HIGHS, MEDS, LOWS)
## Every school x setting gets its own seed, so settings are independent
## samples (needed for the confidence intervals in section 4)
sims   <- run_or_load("policy_grid_indep_seeds",
                      simulate_design(design, independent_seeds = TRUE))

## Runs with an outbreak (> K_MAIN cases, x) out of all runs (n), and the
## probability of an outbreak P = x / n, at every design point
cnt  <- sims[, .(x = sum(size > K_MAIN), n = .N),
             by = .(school, vacc, q_high, q_med, q_low)]
summ <- cnt[, .(school, vacc, q_high, q_med, q_low, P = x / n)]
summ[, vacc_lab := factor(vacc_lab(vacc), levels = VACC_LABS)]
```

`r nrow(design)` quarantine/vaccination settings per school &times;
`r length(SCHOOLS)` schools &times; `r format(N_SIMS, big.mark = ",")` simulations
= `r format(nrow(sims), big.mark = ",")` runs.

::: {.callout-important}
## What "outbreak risk" means in this report

Each simulation starts with one infected student. An **outbreak** is a
simulation that ends with **more than `r K_MAIN` total cases**. The
**outbreak risk** of a setting is the share of its
`r format(N_SIMS, big.mark = ",")` simulations that ended this way, i.e.
P(final outbreak size > `r K_MAIN`). Figures 1 to 5 and the conclusions use
this definition; Figure 6 repeats it for several other thresholds.
:::

## 1. The problem

With minimal quarantine (only high-risk contacts quarantined), how likely is
an outbreak?

```{r fig1, fig.width=10, fig.height=4.5}
d1 <- summ[q_med == 0 & q_low == 0]

p1 <- ggplot(d1, aes(factor(q_high, levels = HIGHS), vacc_lab, fill = P)) +
  geom_tile(colour = "white", linewidth = 1) +
  geom_text(aes(label = pct(P, accuracy = 1), colour = P > 0.5),
            size = 4, show.legend = FALSE) +
  scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black")) +
  scale_fill_gradient(low = "#fff5eb", high = "#b30000", limits = c(0, 1),
                      labels = pct, name = ylab_risk) +
  facet_wrap(~ school) +
  labs(x = "High-risk quarantine (days)", y = "Students vaccinated",
       title = "1. Outbreak risk with only high-risk contacts quarantined",
       subtitle = "Darker = more dangerous") +
  theme(panel.grid = element_blank())

fig(p1, "fig1_baseline_risk.png", 10, 4.5)
```

## 2. What drives outbreak risk?

Each panel changes **one** factor while the others stay fixed, and shows the
outbreak risk before (grey dot) and after (arrow). The fixed values are
named in each panel's title. Vaccination is compared at one fixed quarantine
policy; each quarantine tier is compared at the lowest and highest
vaccination level.

Each risk comes from a finite number of simulations, so two settings can
differ a little just by chance. Arrows are therefore coloured by whether the
change is **clear**: blue if the risk is clearly lower, red if clearly
higher, and grey if the 95% confidence interval of the change includes 0
(the change can't be told apart from simulation noise). The table below the
figure gives the numbers.

```{r fig2a-data}
## The comparisons shown in Figure 2. Every setting used must be in the
## simulated grid (and satisfy high >= medium >= low); otherwise the code
## stops with a message naming the missing setting.
FIG2A <- list(
  vacc   = list(title = "Increasing vaccination",
                from = min(VACCS), to = max(VACCS),
                fixed = c(q_high = 14L, q_med = 7L, q_low = 0L)),
  q_high = list(title = "Increasing high-risk quarantine",
                from = 14L, to = 21L,     # with medium = 14, high can only be 14 or 21
                fixed = c(q_med = 14L, q_low = 0L), vaccs = c(0.60, 0.85)),
  q_med  = list(title = "Increasing medium-risk quarantine",
                from = 0L, to = 14L,
                fixed = c(q_high = 14L, q_low = 0L), vaccs = c(0.60, 0.85)),
  q_low  = list(title = "Increasing low-risk quarantine",
                from = 0L, to = 7L,
                fixed = c(q_high = 14L, q_med = 7L), vaccs = c(0.60, 0.85)))

TIER_NAMES <- c(q_high = "high", q_med = "medium", q_low = "low")

## Outbreak counts (x out of n runs) of one simulated setting
lookup_xn <- function(s, v, h, m, l) {
  r <- cnt[.(s, v, as.integer(h), as.integer(m), as.integer(l)),
           on = .(school, vacc, q_high, q_med, q_low)]
  if (nrow(r) != 1 || is.na(r$x))
    stop(sprintf("Figure 2: %s, %s vaccinated, %d/%d/%d days was not simulated",
                 s, pct(v), h, m, l))
  list(x = r$x, n = r$n)
}

## "Increasing vaccination: 60% \u2192 85% (quarantine 14/7/0 days)"
panel_title <- function(var, sp) {
  change <- if (var == "vacc") sprintf("%s \u2192 %s", pct(sp$from), pct(sp$to))
            else sprintf("%d \u2192 %d days", sp$from, sp$to)
  fixed <- if (var == "vacc")
    sprintf("quarantine %s days", paste(sp$fixed, collapse = "/"))
  else paste(sprintf("%s %d", TIER_NAMES[names(sp$fixed)], sp$fixed), collapse = ", ")
  sprintf("%s: %s  (%s%s)", sp$title, change, fixed, if (var == "vacc") "" else " days")
}

d2a <- rbindlist(lapply(names(FIG2A), function(var) {
  sp <- FIG2A[[var]]
  vs <- if (var == "vacc") NA_real_ else sp$vaccs
  rbindlist(lapply(SCHOOLS, function(s) rbindlist(lapply(vs, function(v) {
    q_at <- function(val) {
      x <- c(sp$fixed, setNames(val, var))
      list(v = if (var == "vacc") val else v,
           h = x[["q_high"]], m = x[["q_med"]], l = x[["q_low"]])
    }
    a <- q_at(sp$from); b <- q_at(sp$to)
    ca <- lookup_xn(s, a$v, a$h, a$m, a$l)
    cb <- lookup_xn(s, b$v, b$h, b$m, b$l)
    data.table(factor = var, school = s,
               label  = if (var == "vacc") s
                        else sprintf("%s \u00b7 %s vaccinated", s, pct(v)),
               x_before = ca$x, n_before = ca$n, x_after = cb$x, n_after = cb$n,
               before = ca$x / ca$n, after = cb$x / cb$n)
  }))))
}))

d2a[, factor_title := factor(vapply(factor, function(v) panel_title(v, FIG2A[[v]]), ""),
                             levels = vapply(names(FIG2A), function(v) panel_title(v, FIG2A[[v]]), ""))]
d2a[, case := factor(paste(factor, label, sep = "|"),
                     levels = rev(unique(paste(factor, label, sep = "|"))))]
## Change = risk after - risk before, with a 95% Newcombe CI
## (settings use separate seeds, so the two risks are independent)
chg <- newcombe_diff(d2a$x_after, d2a$n_after, d2a$x_before, d2a$n_before)
d2a[, `:=`(change = chg$diff, change_lo = chg$lo, change_hi = chg$hi)]
d2a[, direction := factor(fcase(change_hi < 0, "Clearly lower",
                                change_lo > 0, "Clearly higher",
                                default = "No clear change"),
                          levels = c("Clearly lower", "No clear change", "Clearly higher"))]
```

```{r fig2a, fig.width=11, fig.height=10}
p2a <- ggplot(d2a, aes(y = case)) +
  geom_segment(aes(x = before, xend = after, yend = case, colour = direction),
               linewidth = 1.2,
               arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +
  geom_point(aes(x = before), colour = "grey45", size = 3) +
  ## labels sit outside the arrow ends so they never overlap
  geom_text(aes(x = before, label = pct(before, accuracy = 1),
                hjust = ifelse(after <= before, -0.4, 1.4)),
            colour = "grey30", size = 3.3) +
  geom_text(aes(x = after, label = pct(after, accuracy = 1), colour = direction,
                hjust = ifelse(after <= before, 1.3, -0.3)),
            size = 3.3, fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = c("Clearly lower"   = "#2171b5",
                                 "No clear change" = "grey60",
                                 "Clearly higher"  = "#b30000"),
                      name = "Change in risk (95% CI)", drop = FALSE) +
  scale_x_continuous(labels = pct, limits = c(0, 1)) +
  scale_y_discrete(labels = function(x) sub("^[^|]*\\|", "", x)) +
  facet_wrap(~ factor_title, ncol = 1, scales = "free_y") +
  labs(x = ylab_risk, y = NULL,
       title = "2. What happens when one factor changes?",
       subtitle = "Grey dot = before, arrow = after. Everything not named in the panel title stays fixed.\nGrey arrow = change within simulation noise (95% CI includes 0)") +
  theme(panel.grid.major.y = element_blank(),
        strip.text  = element_text(hjust = 0, size = 11),
        axis.text.y = element_text(size = 9.5))

fig(p2a, "fig2_factor_effects.png", 11, 10)
```

**Small benefits, with 95% confidence intervals.** Large changes are clear
from the figure, so this table only lists the comparisons where the
**benefit is under `r round(100 * FIG2_TABLE_BENEFIT)` percentage points**,
where simulation noise matters most. Benefit = risk before &minus; risk after,
in percentage points: positive = lower risk, negative = higher risk.
*Clearly lower / higher* means the whole interval is above / below 0; *No
clear change* means the interval includes 0. Intervals reflect simulation
noise only, not uncertainty in the model's assumptions.

```{r fig2-table, results='asis'}
pts_signed <- function(v) sprintf("%+.1f", 100 * v)

## Benefit = before - after (the change with its sign flipped)
d2a[, `:=`(benefit = -change, benefit_lo = -change_hi, benefit_hi = -change_lo)]
small <- d2a[benefit < FIG2_TABLE_BENEFIT]

if (nrow(small) == 0) {
  cat(sprintf("\n\nEvery comparison has a benefit of %s points or more.\n\n",
              round(100 * FIG2_TABLE_BENEFIT)))
} else {
  for (v in names(FIG2A)) {
    d <- small[factor == v]
    cat(sprintf("\n\n*%s*\n\n", panel_title(v, FIG2A[[v]])))
    if (nrow(d) == 0) {
      cat(sprintf("All comparisons in this panel have a benefit of %s points or more.\n\n",
                  round(100 * FIG2_TABLE_BENEFIT)))
      next
    }
    print(knitr::kable(
      d[, .(Setting = label,
            `Risk before` = pct(before, accuracy = 0.1),
            `Risk after`  = pct(after,  accuracy = 0.1),
            `Benefit, points (95% CI)` = sprintf("%s (%s to %s)", pts_signed(benefit),
                                                 pts_signed(benefit_lo), pts_signed(benefit_hi)),
            Verdict = as.character(direction))],
      align = c("l", "r", "r", "r", "l")))
  }
}
```

## 3. Which quarantine combinations keep risk low?

Every quarantine combination, coloured by its outbreak risk on a continuous
scale: **light = low risk, dark = high risk**. The number in each square is
its risk. Squares with a **black outline** have a risk of
`r pct(FIG3_HIGHLIGHT)` or less. Columns are medium / low-risk days, grouped
by high-risk days; low-risk days are shown in weekly steps to keep the figure
readable.

```{r fig3, fig.width=11, fig.height=7}
## Shortest policy per target (used in the conclusions)
d3 <- rbindlist(lapply(TARGETS, function(tg) {
  summ[, {
    ok <- .SD[P <= tg][order(q_high + q_med + q_low, q_low, q_med)]
    if (nrow(ok) == 0) .(days = NA_integer_, label = "Not possible")
    else .(days  = ok$q_high[1] + ok$q_med[1] + ok$q_low[1],
           label = sprintf("%d / %d / %d", ok$q_high[1], ok$q_med[1], ok$q_low[1]))
  }, by = .(school, vacc_lab)][, target := tg]
}))

d3p <- summ[q_low %in% FIG3_LOWS]
ML_LEVELS <- unique(d3p[order(q_med, q_low), sprintf("%d/%d", q_med, q_low)])
d3p[, ml := factor(sprintf("%d/%d", q_med, q_low), levels = ML_LEVELS)]
d3p[, high_lab := factor(sprintf("High-risk %d days", q_high),
                         levels = sprintf("High-risk %d days", HIGHS))]

fig3_max <- if (is.na(FIG3_MAX)) max(d3p$P) else FIG3_MAX
d3p[, dark := P > 0.55 * fig3_max]     # white text on dark squares

p3 <- ggplot(d3p, aes(ml, vacc_lab, fill = P)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  scale_fill_gradientn(colours = FIG3_COLORS, limits = c(0, fig3_max),
                       oob = scales::squish, labels = pct,
                       breaks = sort(unique(c(TARGETS[TARGETS < fig3_max],
                                              pretty(c(0, fig3_max), 4)))),
                       name = "Outbreak risk") +
  facet_grid(school ~ high_lab, scales = "free_x", space = "free_x") +
  labs(x = "Medium / low-risk quarantine days", y = "Students vaccinated",
       title = "3. Every quarantine combination: which ones keep risk low?",
       subtitle = paste0("Each square is one policy. Light = low risk, dark = high risk.",
                         if (!is.na(FIG3_HIGHLIGHT))
                           sprintf(" Black outline = risk %s or less.", pct(FIG3_HIGHLIGHT)) else "",
                         "\nRead left to right to find the shortest policy with a low risk")) +
  guides(fill = guide_colourbar(barwidth = unit(8, "cm"))) +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 9))

if (FIG3_LABELS)
  p3 <- p3 + geom_text(aes(label = pct(P, accuracy = 1), colour = dark),
                       size = 2.7, show.legend = FALSE) +
    scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black"))

if (!is.na(FIG3_HIGHLIGHT))
  p3 <- p3 + geom_tile(data = d3p[P <= FIG3_HIGHLIGHT], fill = NA,
                       colour = "black", linewidth = 0.8,
                       width = 0.92, height = 0.92)

fig(p3, "fig3_policy_options.png", 11, 7)
```

## 4. How many low-risk quarantine days are worth it?

For each fixed high/medium policy: the fewest low-risk days that give
`r pct(OPT_SHARE)` of the benefit low-risk quarantine can provide. Each cell
shows those days and the risk without &rarr; with them.

```{r fig4, fig.width=12, fig.height=5.8}
d4 <- summ[q_med > 0, {
  o     <- .SD[order(q_low)]
  risk0 <- o$P[o$q_low == 0]
  gain  <- cummax(risk0 - o$P)     # more days never counted as worse (smooths noise)
  G     <- max(gain)
  best  <- if (G < MIN_GAIN - 1e-9) 0L else o$q_low[which(gain >= OPT_SHARE * G)[1]]   # 1e-9: rounding
  ## low-risk days giving the largest benefit (the longest allowed if none helps)
  max_low <- if (G > 0) o$q_low[which.max(gain)] else max(o$q_low)
  .(opt_low  = best,
    max_low  = max_low,
    risk0    = risk0,
    risk_opt = risk0 - gain[o$q_low == best],
    risk_max = risk0 - G)
}, by = .(school, vacc, vacc_lab, q_high, q_med)]

hm_levels <- unique(d4[order(-q_high, -q_med), sprintf("High %d \u00b7 Med %d", q_high, q_med)])
d4[, hm := factor(sprintf("High %d \u00b7 Med %d", q_high, q_med), levels = rev(hm_levels))]

p4 <- ggplot(d4, aes(vacc_lab, hm, fill = opt_low)) +
  geom_tile(colour = "white", linewidth = 1) +
  geom_text(aes(label = ifelse(opt_low == 0, "Not\nworth it",
                               sprintf("%d days\n%s \u2192 %s", opt_low,
                                       pct(risk0, accuracy = 1), pct(risk_opt, accuracy = 1))),
                colour = opt_low > 10),
            size = 2.8, lineheight = 0.9, show.legend = FALSE) +
  scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black")) +
  scale_fill_gradient(low = "#f7fbff", high = "#08519c", limits = c(0, max(LOWS)),
                      name = "Optimal low-risk days") +
  facet_wrap(~ school) +
  labs(x = "Students vaccinated", y = "Fixed policy (days)",
       title = "4. How many low-risk quarantine days are worth it?",
       subtitle = sprintf(
         "Each cell: best number of low-risk days, and outbreak risk without \u2192 with them. \"Not worth it\" = risk drops less than %s",
         pct(MIN_GAIN))) +
  theme(panel.grid = element_blank())

fig(p4, "fig4_optimal_low.png", 12, 5.8)
```

### How sure is "not worth it"?

A cell is *not worth it* when the largest benefit low-risk quarantine gives
is under `r round(100 * MIN_GAIN)` percentage points. The tables below list
only those cells, school by school. For each, the risk with 0 low-risk days
is compared with the risk at the number of low-risk days that gave the
largest benefit (or the longest allowed, if none helped), with a **95%
confidence interval** for the benefit (Newcombe method for a difference of
two proportions; every setting has its own seed, so the two risks are
independent).

- **Benefit** = risk with 0 low-risk days &minus; risk with those days, in
  percentage points (positive = lower risk).
- **Clearly under `r round(100 * MIN_GAIN)` points?** *Yes* if the whole
  interval is below `r round(100 * MIN_GAIN)` points, so the true benefit is
  small. *No* if the interval reaches `r round(100 * MIN_GAIN)` points or
  more: the benefit is probably small, but these simulations can't rule out
  a benefit of that size.

The intervals show simulation uncertainty only, not uncertainty in the
model's assumptions.

```{r fig4-table-data}
## Only the "Not worth it" cells
t4 <- d4[opt_low == 0]
t4 <- cnt[q_low == 0, .(school, vacc, q_high, q_med, x0 = x, n0 = n)][
  t4, on = .(school, vacc, q_high, q_med)]
t4 <- cnt[, .(school, vacc, q_high, q_med, max_low = q_low, x1 = x, n1 = n)][
  t4, on = .(school, vacc, q_high, q_med, max_low)]

## Benefit = risk with 0 low-risk days - risk with max_low days
ben <- newcombe_diff(t4$x0, t4$n0, t4$x1, t4$n1)
t4[, `:=`(ben = ben$diff, ben_lo = ben$lo, ben_hi = ben$hi)]

pts <- function(v) sprintf("%+.1f", 100 * v)
t4[, `:=`(
  Vaccinated              = as.character(vacc_lab),
  `High / medium days`    = sprintf("%d / %d", q_high, q_med),
  `Low-risk days compared` = sprintf("0 vs %d", max_low),
  `Risk, 0 days`          = pct(x0 / n0, accuracy = 0.1),
  `Risk, with days`       = pct(x1 / n1, accuracy = 0.1),
  `Benefit, points (95% CI)` = sprintf("%s (%s to %s)", pts(ben), pts(ben_lo), pts(ben_hi)),
  `Clearly under limit?`  = ifelse(ben_hi < MIN_GAIN, "Yes", "No"))]
setorder(t4, school, vacc, -q_high, -q_med)

T4_COLS <- c("Vaccinated", "High / medium days", "Low-risk days compared",
             "Risk, 0 days", "Risk, with days",
             "Benefit, points (95% CI)", "Clearly under limit?")
```

::: {.panel-tabset}

```{r fig4-table, results='asis'}
for (s in SCHOOLS) {
  cat("\n\n#### ", s, "\n\n")
  d <- t4[school == s]
  if (nrow(d) == 0) {
    cat("No *not worth it* cells: low-risk quarantine lowers the risk by at least",
        round(100 * MIN_GAIN), "points in every setting for this school.\n\n")
  } else {
    print(knitr::kable(d[, ..T4_COLS], align = c("l", "l", "l", "r", "r", "r", "c"),
                       col.names = sub("limit", paste(round(100 * MIN_GAIN), "points"),
                                       T4_COLS)))
  }
  cat("\n\n")
}
```

:::

## 5. How big do outbreaks get?

```{r fig5-helpers}
SIZE_LABS <- c("10 or fewer", "11\u201335", "36\u201350", "More than 50")
SIZE_COLS <- c("#fee5d9", "#fcae91", "#fb6a4a", "#a50f15")

## Share of runs in each size band, for the policy chosen per school/vacc
size_shares <- function(policy) {
  d <- sims[policy, on = .(school, vacc, q_high, q_med, q_low)]
  d[, size_cat := cut(size, breaks = c(-Inf, 10, 35, 50, Inf), labels = SIZE_LABS)]
  d <- d[, .N, by = .(school, vacc, q_high, q_med, q_low, size_cat)][
    , share := N / sum(N), by = .(school, vacc, q_high, q_med, q_low)]
  d[, vacc_lab := factor(vacc_lab(vacc), levels = VACC_LABS)]
  d[]
}

## Optimal policy per school and vaccination (explained in 5b)
opt_pick <- function(dt) {
  dt    <- copy(dt)
  risk0 <- dt[q_high == min(HIGHS) & q_med == 0 & q_low == 0, P]
  best  <- min(dt$P)
  dt[, days := q_high + q_med + q_low]
  need <- if (risk0 - best < MIN_GAIN - 1e-9) 0 else OPT_SHARE * (risk0 - best)
  dt[risk0 - P >= need - 1e-12][order(days, P)][1, .(q_high, q_med, q_low, P,
                                                     P0 = risk0, Pmin = best)]
}

opt_pol <- summ[, opt_pick(.SD), by = .(school, vacc)]
opt_pol[, policy := sprintf("%d/%d/%d", q_high, q_med, q_low)]

## A worked example for the explanation in 5b: the first school and
## vaccination level where quarantine is worth it
ex <- opt_pol[P0 - Pmin >= MIN_GAIN][1]
if (is.na(ex$school)) ex <- opt_pol[1]      # quarantine never worth it anywhere
ex_G      <- ex$P0 - ex$Pmin
ex_target <- ex$P0 - OPT_SHARE * ex_G

size_plot <- function(d, facet = facet_wrap(~ school)) {
  ggplot(d, aes(vacc_lab, share, fill = size_cat)) +
    geom_col(width = 0.7, position = position_stack(reverse = TRUE)) +
    scale_fill_manual(values = SIZE_COLS, name = "Total cases", drop = FALSE) +
    facet +
    labs(x = "Students vaccinated", y = "Share of simulations") +
    theme(panel.grid.major.x = element_blank())
}
```

::: {.panel-tabset}

### 5a. Only high-risk contacts quarantined

```{r fig5a, fig.width=10, fig.height=8.5}
## Only high-risk contacts quarantined: 7/0/0, 14/0/0 and 21/0/0
hi_pol <- unique(summ[q_med == 0 & q_low == 0, .(school, vacc, q_high, q_med, q_low)])

d5a <- size_shares(hi_pol)
d5a[, policy_lab := factor(sprintf("High-risk %d days (%d/0/0)", q_high, q_high),
                           levels = sprintf("High-risk %d days (%d/0/0)", HIGHS, HIGHS))]

p5a <- size_plot(d5a, facet_grid(policy_lab ~ school)) +
  scale_y_continuous(labels = pct) +
  labs(title = "5a. How big outbreaks get when only high-risk contacts are quarantined",
       subtitle = "Rows = high-risk quarantine days (medium and low = 0). Light = died out, dark = spread widely")

fig(p5a, "fig5a_outbreak_size_high_only.png", 10, 8.5)
```

### 5b. Optimal quarantine

**How the optimal policy is computed.** For one school and one vaccination
level, let $P(h, m, l)$ be the outbreak risk with $h$ high-, $m$ medium- and
$l$ low-risk quarantine days: the share of simulations with more than
`r K_MAIN` cases. Only policies with $h \ge m \ge l$ are considered.

1. **Risk with minimal quarantine:** $P_0 = P(`r min(HIGHS)`, 0, 0)$.
2. **Lowest risk any tested policy reaches:** $P_{\min} = \min_{h, m, l} P(h, m, l)$.
3. **Largest possible reduction:** $G = P_0 - P_{\min}$.
4. **If quarantine barely helps** ($G <$ `r pct(MIN_GAIN)`), the optimal
   policy is minimal quarantine, `r MIN_POLICY`.
5. **Otherwise, a policy is good enough** if it gets at least
   `r pct(OPT_SHARE)` of the largest reduction:
   $$P_0 - P(h, m, l) \ge `r OPT_SHARE` \times G$$
6. **The optimal policy** is the good-enough policy with the fewest total
   quarantine days $h + m + l$. If several tie, the one with the lowest risk
   is chosen.

*Example:* `r ex$school` school, `r pct(ex$vacc)` vaccinated.
$P_0$ = `r pct(ex$P0, accuracy = 0.1)` and $P_{\min}$ =
`r pct(ex$Pmin, accuracy = 0.1)`, so $G$ = `r sprintf("%.1f", 100 * ex_G)`
percentage points. A good-enough policy must bring the risk down to
`r pct(ex_target, accuracy = 0.1)` or lower
(`r pct(ex$P0, accuracy = 0.1)` &minus; `r OPT_SHARE` &times;
`r sprintf("%.1f", 100 * ex_G)` points). The shortest such policy is
**`r ex$policy`** (`r ex$q_high + ex$q_med + ex$q_low` days in total), with
a risk of `r pct(ex$P, accuracy = 0.1)`.

```{r fig5b, fig.width=10, fig.height=5}
d5b     <- size_shares(opt_pol[, .(school, vacc, q_high, q_med, q_low)])
pol_lab <- opt_pol[, .(school, vacc_lab = factor(vacc_lab(vacc), levels = VACC_LABS), policy)]

p5b <- size_plot(d5b) +
  geom_text(data = pol_lab, aes(vacc_lab, 1, label = policy), inherit.aes = FALSE,
            vjust = -0.6, size = 3.3, fontface = "bold") +
  scale_y_continuous(labels = pct, limits = c(0, 1.08), breaks = seq(0, 1, 0.25)) +
  labs(title = "5b. How big outbreaks get with the optimal quarantine",
       subtitle = sprintf(paste0("Above each bar: optimal quarantine days (high/medium/low), the shortest policy\n",
                                 "giving %s of the largest possible risk reduction. Light = died out, dark = spread widely"),
                          pct(OPT_SHARE)))

fig(p5b, "fig5b_outbreak_size_optimal.png", 10, 5)
```



:::

## Supplement: every quarantine combination, per school

::: {.panel-tabset}

```{r supplement, results='asis', fig.width=11, fig.height=10}
summ[, high_lab := factor(paste0("High-risk = ", q_high, " days"),
                          levels = paste0("High-risk = ", HIGHS, " days"))]

for (s in SCHOOLS) {
  cat("\n\n### ", s, "\n\n")
  ps <- ggplot(summ[school == s],
               aes(factor(q_low, levels = LOWS), factor(q_med, levels = MEDS), fill = P)) +
    geom_tile(colour = "white", linewidth = 0.6) +
    geom_text(aes(label = pct(P, accuracy = 1), colour = P > 0.5),
              size = 2.6, show.legend = FALSE) +
    scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black")) +
    scale_fill_gradient(low = "#fff5eb", high = "#b30000", limits = c(0, 1),
                        labels = pct, name = ylab_risk) +
    facet_grid(paste(vacc_lab, "vaccinated") ~ high_lab) +
    labs(x = "Low-risk quarantine (days)", y = "Medium-risk quarantine (days)",
         title = sprintf("Supplement: %s school, every quarantine combination", s),
         subtitle = "Empty cells are not allowed (high \u2265 medium \u2265 low)") +
    theme(panel.grid = element_blank())
  print(fig(ps, sprintf("supp_grid_%s.png", tolower(s)), 11, 10))
  cat("\n\n")
}
```

:::

## Conclusions

```{r conclusions}
out   <- character()
add   <- function(...) out <<- c(out, sprintf(...))
ascii <- function(x) gsub("\u00b7", "-", gsub("\u2192", "->", x))   # plain-text file

add("CONCLUSIONS  (outbreak = more than %d cases)", K_MAIN)
add("")

## 1. Riskiest school
base <- summ[q_high == min(HIGHS) & q_med == 0 & q_low == 0,
             .(P = mean(P)), by = school][order(-P)]
add("1. THE PROBLEM")
add("   With minimal quarantine, %s school is the riskiest (average %s)",
    as.character(base$school[1]), pct(base$P[1], 1))
add("   and %s school the least risky (average %s).",
    as.character(last(base$school)), pct(last(base$P), 1))
add("")

## 2. Levers
add("2. WHAT MATTERS MOST")
for (v in names(FIG2A)) {
  add("   %s:", ascii(panel_title(v, FIG2A[[v]])))
  for (k in seq_len(nrow(d2a[factor == v]))) {
    r <- d2a[factor == v][k]
    add("     %s: risk %s -> %s (change %s points, 95%% CI %s to %s: %s)",
        ascii(r$label), pct(r$before, 1), pct(r$after, 1),
        pts_signed(r$change), pts_signed(r$change_lo), pts_signed(r$change_hi),
        tolower(r$direction))
  }
}
add("")

## 3. Policy
add("3. WHAT A SCHOOL SHOULD DO")
for (tg in TARGETS) {
  add("   To keep risk under %s:", pct(tg))
  for (s in SCHOOLS) {
    t  <- d3[school == s & target == tg][order(vacc_lab)]
    ok <- t[!is.na(days)]
    mn <- t[label == MIN_POLICY]
    add("     %s: %s; minimal quarantine is enough %s.", s,
        if (nrow(ok)) sprintf("achievable from %s vaccination", as.character(ok$vacc_lab[1]))
        else "not achievable at any tested vaccination level",
        if (nrow(mn)) sprintf("from %s vaccination", as.character(mn$vacc_lab[1]))
        else "at no tested level")
  }
}
add("")

## 4. Optimal low-risk days
add("4. LOW-RISK QUARANTINE")
for (s in SCHOOLS) {
  o <- d4[school == s]
  add("   %s: the best low-risk quarantine is usually about %g days.", s, median(o$opt_low))
  add("     It is not worth it (risk drops less than %s) in %s of high/medium settings.",
      pct(MIN_GAIN), pct(mean(o$opt_low == 0), 1))
  add("     On average, risk goes %s -> %s with the best number of days,",
      pct(mean(o$risk0), 1), pct(mean(o$risk_opt), 1))
  add("     and only to %s even with the longest low-risk quarantine.", pct(mean(o$risk_max), 1))
}
add("")

## 5. Size
add("5. SIZE OF OUTBREAKS")
for (s in SCHOOLS) {
  b <- sims[school == s & q_high == min(HIGHS) & q_med == 0 & q_low == 0,
            .(share = mean(size > 50)), by = vacc][order(vacc)]
  add("   %s, minimal quarantine: chance of more than 50 cases falls from %s at %s to %s at %s vaccination.",
      s, pct(b$share[1], 1), pct(b$vacc[1]), pct(last(b$share), 1), pct(last(b$vacc)))

  op <- opt_pol[school == s][order(vacc)]
  bo <- sims[op, on = .(school, vacc, q_high, q_med, q_low)][
    , .(share = mean(size > 50)), by = vacc][order(vacc)]
  add("     Optimal quarantine (high/medium/low days): %s.",
      paste(sprintf("%s at %s", op$policy, pct(op$vacc)), collapse = "; "))
  add("     With it, the chance of more than 50 cases is %s at %s and %s at %s vaccination.",
      pct(bo$share[1], 1), pct(bo$vacc[1]), pct(last(bo$share), 1), pct(last(bo$vacc)))
}

writeLines(out, file.path(RESULTS_DIR, "conclusions.txt"))
```

```{r print-conclusions, results='asis'}
cat("```text", out, "```", sep = "\n")
```

