---
format: gfm
title: Measles Tiered Quarantine Simulation
author: George G. Vega Yon, Ph.D.
date: 2025-10-23
toc: true
toc-location: body
---

```{=markdown}
[![](https://github.com/EpiForeSITE/software/raw/e82ed88f75e0fe5c0a1a3b38c2b94509f122019c/docs/assets/foresite-software-badge.svg)](https://github.com/EpiForeSITE/software)

> [!CAUTION]
> This project is a work in progress. Use it at your own risk. **This model simulates a single school, so community transmission is not included**.

> [!IMPORTANT]
> The model makes several assumptions that may not hold in real-world scenarios. One important assumption is that interactions between agents are based solely on class assignments, and not based on friendship networks or other social structures. This assumption reflects strongly in the effect of mid-risk quarantine (see below for more details).
```

## Description of the model

We are using the `ModelMeaslesMixingRiskQuarantine()` model from the `epiworldR` package (version 0.10.0.0 or higher). This model uses a mixing matrix to represent how agents' interactions are distributed. In this case, groups represent classrooms in a school, so kids will have more interactions with kids in their own class than with kids from other classes.

The model contains the full disease progression for measles, including the incubation period, prodromal phase, and rash phase. The model assumes that kids in the rash phase are isolated from their peers. Nonetheless, the model is calibrated to reflect the measles $\mathcal{R}_0$ of 15, so kids are infectious during the prodromal period. Agents can also become hospitalized.

The quarantine process is triggered when an agent is detected during the rash period. This could occur for more than one agent in the same step. Quarantine process only applies to unvaccinated agents, and the duration of quarantine depends on the risk level of the agent, as defined below.

## Setup

This document illustrates an experiment using the `epiworldR` package to simulate measles transmission under a tiered quarantine strategy. In the tier quarantine system, agents have different quarantine durations as a function of their risk level, which are defined as follows:

- **High risk**: Agents in the same classroom as an infected individual.
- **Medium risk**: Agents who were not in the same classroom as the infected individual, but were in direct contact with them.
- **Low risk**: Agents who were not in the same classroom or direct contact with the infected individual.

Quarantine only applies to unvaccinated agents. The simulation settings are as follows:

- Individual school with 600 students distributed across 20 classes (30 students per class).
- The contact rate is given by our previous estimates for within-class and between-class interactions: 83% of contacts occur within the same class, while 17% occur between different classes.
- The basic reproduction number (R0) is set to 15, reflecting the high transmissibility of measles.
- Contact tracing is assumed to be 100% effective, as agents' willingness to isolate and quarantine.

The following code block sets up some of the simulation parameters, including sourcing the simulator function in the file [`simulator.R`](./simulator.R):

```{r}
#| label: setup-paramters
library(epiworldR)
library(data.table)
library(ggplot2)
source("simulator.R")

# Simulation parameters
n_sims    <- 2000
n_agents  <- 600
n_classes <- 20
n_agents_per_class <- n_agents / n_classes
n_days    <- 100
n_threads <- 10

# Makesure it's even
stopifnot(n_agents_per_class %% 1 == 0)
```

The particular disease parameters for measles, including the mixing matrix that will be used for the simulation, are defined as follows:

```{r}
#| label: disease-parameters
# Disease parameters
R0 <- 15
contact_rate <- 20
incubation <- 12
prodromal   <- 4
rash        <- 3

# Creating the mixing matrix
within_class_contact_rate <- 0.83
between_class_contact_rate <- 1 - within_class_contact_rate

contact_matrix <- matrix(
  between_class_contact_rate / (n_classes - 1),
  nrow = n_classes,
  ncol = n_classes
)
diag(contact_matrix) <- within_class_contact_rate

# Calibrating infection probability
p_infect <- R0 / (contact_rate) * (1/prodromal)
```

We will test the model using the following scenarios:

- 50%, 80%, and 95% vaccination coverage.
- Quarantine days set to 0, 7, 14, and 21 days.

To assess the risk effect between different quarantine strategies, we report the probability of observing outbreaks of various sizes rather than means and medians. Specifically, we calculate:

- **P(≥10)**: Probability that an outbreak reaches 10 or more infected individuals
- **P(≥20)**: Probability that an outbreak reaches 20 or more infected individuals
- **P(≥50)**: Probability that an outbreak reaches 50 or more infected individuals

This approach provides a clearer picture of the distribution's tail behavior and allows for better comparison of risk across different strategies.

## Parameters & references

The table below lists every parameter of `ModelMeaslesMixingRiskQuarantine()` used in this analysis, including the ones left at their default value. **Value used** is what the simulations in this document ran with. **Package default** is the default in the [`measles`](https://github.com/UofUEpiBio/measles) R package, which now hosts this model and keeps the canonical, cited parameter table: [`measles_parameters.csv`](https://github.com/UofUEpiBio/measles/blob/main/inst/extdata/measles_parameters.csv) (also available as `measles::measles_parameters()`; see the [parameters vignette](https://github.com/UofUEpiBio/measles/blob/main/vignettes/parameters.qmd)). Sources follow that table unless the Notes say otherwise.

The results in this document were produced with `epiworldR` 0.10.0.0, where the model still lived. Two of that version's defaults differ from the current `measles` package: vaccine efficacy (0.99 vs 0.97) and the contact-tracing argument name (`contact_tracing_days_prior` vs `contact_tracing_days_window`). `epiworldR` 0.10.0.0 also had a `contact_rate` argument that the `measles` package removed; there, the contact matrix holds the expected number of contacts instead.

| Parameter | Value used | Package default | Source | Notes / why different |
|---|---|---|---|---|
| R0 (target) | 15 | 15 | Guerra et al. 2017, *Lancet Infect Dis* 17(12):e420–e428, [doi:10.1016/S1473-3099(17)30307-9](https://doi.org/10.1016/S1473-3099(17)30307-9) | Not a model argument. Used to calibrate the transmission probability (see below). 15 is the midpoint of the 12–18 range. |
| Contact rate (`contact_rate`) | 20 contacts/day | Not an argument in `measles` (scale `contact_matrix` instead). `epiworldR` 0.10.0.0 default: `15 / transmission_rate / prodromal_period` | Team assumption | **Differs.** Not passed to the model explicitly: the `epiworldR` 0.10.0.0 default formula gives 15 / 0.1875 / 4 = 20, the same value used in the calibration. No source is recorded for 20 contacts/day. |
| Transmission probability (`transmission_rate`) | 0.1875 | 0.9 | Derived: R0 / (contact rate × prodromal period) = 15 / (20 × 4) | **Differs.** This analysis fixes the contact rate at 20 and calibrates the transmission probability to R0 = 15. The package fixes transmission high (0.9) and calibrates contacts instead. Both target the same R0. |
| Contact matrix (`contact_matrix`) | 20 × 20 classes; 0.83 within class, 0.17 / 19 to each other class | No default | Team estimate: "our previous estimates for within-class and between-class interactions" | Row-stochastic mixing proportions (`epiworldR` 0.10.0.0 semantics), multiplied by the contact rate of 20. In the `measles` package the equivalent input is 20 × this matrix (expected contacts per day). No citation is recorded for the 83% / 17% split. |
| Population size (`n`) | 600 (20 classes of 30) | No default | Scenario-specific input | A single school. |
| Prevalence (`prevalence`) | 1 / 600 (one index case) | No default | Scenario-specific input | |
| Vaccination coverage (`prop_vaccinated`) | 0, 0.5, 0.8, 0.9 (by scenario) | No default | Scenario-specific input | Scenarios, not estimates. |
| Vaccine efficacy (`vax_efficacy`) | 0.99 (`epiworldR` 0.10.0.0 default) | 0.97 | Liu et al. 2015, *BMC Public Health* 15:447, [doi:10.1186/s12889-015-1766-6](https://doi.org/10.1186/s12889-015-1766-6) (0.99); current default 0.97 from the [Utah DHHS Measles Disease Plan](https://epi.utah.gov/wp-content/uploads/Measles-disease-plan.pdf) ("~97%") and CDC | **Differs.** Not passed explicitly, so the results used the `epiworldR` 0.10.0.0 default of 0.99. Re-running with the `measles` package would use 0.97. |
| Incubation period (`incubation_period`) | 12 days (default) | 12 | [Utah DHHS Measles Disease Plan](https://epi.utah.gov/wp-content/uploads/Measles-disease-plan.pdf): exposure to prodrome averages 8–12 days | Defined as `incubation <- 12` in the setup code but not passed; the model default is the same. |
| Prodromal period (`prodromal_period`) | 4 days (default) | 4 | [Utah DHHS Measles Disease Plan](https://epi.utah.gov/wp-content/uploads/Measles-disease-plan.pdf): prodrome 2–4 days; contagious 4 days before rash onset | Not passed; the same value (4) is used in the transmission calibration. |
| Rash period (`rash_period`) | 3 days (default) | 3 | [Utah DHHS Measles Disease Plan](https://epi.utah.gov/wp-content/uploads/Measles-disease-plan.pdf): contagious to 4 days after rash onset; infectivity minimal after day 2 of rash | Defined as `rash <- 3` but not passed; the model default is the same. Only the infectious part of the rash is modeled. |
| Hospitalization rate (`hospitalization_rate`) | 0.2 per day (default) | 0.2 | Conservative value agreed with Utah DHHS; recent analyses use 10% per Jones et al. 2026, *NEJM Evid* 5(8), [doi:10.1056/EVIDpha2600141](https://doi.org/10.1056/EVIDpha2600141) | A daily **rate**, not a probability: p = h / (h + 1/rash) = 0.2 / (0.2 + 1/3) ≈ 0.375. |
| Hospitalization period (`hospitalization_period`) | 7 days (default) | 7 | Assumption | Observed stays are shorter (Utah mean 2.1 nights, Jones et al. 2026). |
| Days undetected (`days_undetected`) | 2 days (default) | 2 | Assumption: about 2 days from active case to public-health notification | |
| Isolation period (`isolation_period`) | 4 days (default) | 4 | [Utah DHHS Measles Disease Plan](https://epi.utah.gov/wp-content/uploads/Measles-disease-plan.pdf): isolate until 4 days after rash onset | |
| Isolation willingness (`isolation_willingness`) | 1 (default) | 1 | Assumption (field experience) | Perfect compliance, as stated in the Setup. |
| Quarantine willingness (`quarantine_willingness`) | 1 (default) | 1 | Assumption (field experience) | Perfect compliance, as stated in the Setup. |
| Contact-tracing success rate (`contact_tracing_success_rate`) | 1 (default) | 1 | Assumption (field experience) | 100% effective contact tracing, as stated in the Setup. |
| Contact-tracing window (`contact_tracing_days_prior` in `epiworldR` 0.10.0.0; `contact_tracing_days_window` in `measles`) | 7 days | 4 | Default 4 matches the Utah DHHS exposure definition (4 days before through 4 days after rash onset) | **Differs.** Set explicitly to 7 in `simulator.R`. The rationale for 7 is not recorded. |
| Detection rate in quarantine (`detection_rate_quarantine`) | 0 | 0.5 | Assumption | **Differs.** Set explicitly to 0 in `simulator.R`: prodromal agents in quarantine are never detected, so detection happens only through the rash. The rationale is not recorded; the package default (0.5) also has no recorded source. |
| Quarantine period, high risk (`quarantine_period_high`) | 21 days in most scenarios; 0 or 14 in some | 21 | Our own experiments and discussions with Utah DHHS | **Experimental; no published source.** Set explicitly per scenario. The 21-day tier matches the Utah DHHS standard quarantine (21 days since last exposure). |
| Quarantine period, medium risk (`quarantine_period_medium`) | 0, 7, 10, 14 or 21 days (by scenario) | 14 | Our own experiments and discussions with Utah DHHS | **Experimental; no published source.** Set explicitly per scenario; these lengths are what the experiment varies. |
| Quarantine period, low risk (`quarantine_period_low`) | 0, 7, 10, 14 or 21 days (by scenario) | 7 | Our own experiments and discussions with Utah DHHS | **Experimental; no published source.** Set explicitly per scenario; these lengths are what the experiment varies. |

## Scenarios

### Scenario: No vaccination, only one risk level quarantined

```{r}
#| label: no-vaccination-single-risk
ans_none        <- simulator(duration = c(0L, 0L, 0L), vaccinated = 0.0)
ans_only_high   <- simulator(duration = c(21L, 0L, 0L), vaccinated = 0.0)
ans_only_medium <- simulator(duration = c(0L, 21L, 0L), vaccinated = 0.0)
ans_only_low    <- simulator(duration = c(0L, 0L, 21L), vaccinated = 0.0)

# Tabulating the results
tabulator(
  list(
    "No Quarantine" = ans_none,
    "Only High Risk Quarantine" = ans_only_high,
    "Only Medium Risk Quarantine" = ans_only_medium,
    "Only Low Risk Quarantine" = ans_only_low
  )
)
```

### Scenario: 50% vaccination, tiered quarantine

For this scenario, we simulate with 50% vaccination coverage and compare the following tiered quarantine strategies:

- Baseline: 21 days for all risk levels.
- Strategy 1: 21 days for high risk, 14 days for medium and low risk.
- Strategy 2: 21 days for high risk, 7 days for medium and low risk.
- Strategy 3: 21 days for high risk, no quarantine for medium and low risk.

```{r}
#| label: vaccination-50-tiered
ans_50_baseline <- simulator(duration = c(21L, 21L, 21L), vaccinated = 0.5)
ans_50_strategy1 <- simulator(duration = c(21L, 14L, 14L), vaccinated = 0.5)
ans_50_strategy2 <- simulator(duration = c(21L, 7L, 7L), vaccinated = 0.5)
ans_50_strategy3 <- simulator(duration = c(21L, 0L, 0L), vaccinated = 0.5)

# Tabulating the results
tabulator(
  list(
    "Baseline (21,21,21)" = ans_50_baseline,
    "Strategy 1 (21,14,14)" =  ans_50_strategy1,
    "Strategy 2 (21,7,7)" = ans_50_strategy2,
    "Strategy 3 (21,0,0)" = ans_50_strategy3
  )
)
```

### Scenario: 80% vaccination, tiered quarantine

Similar to the previous scenario, we simulate with 80% vaccination coverage and compare the same tiered quarantine strategies:

```{r}
#| label: vaccination-80-tiered
ans_80_baseline <- simulator(duration = c(21L, 21L, 21L), vaccinated = 0.8)
ans_80_strategy1 <- simulator(duration = c(21L, 14L, 14L), vaccinated = 0.8)
ans_80_strategy2 <- simulator(duration = c(21L, 7L, 7L), vaccinated = 0.8)
ans_80_strategy3 <- simulator(duration = c(21L, 0L, 0L), vaccinated = 0.8)

# Tabulating the results
tabulator(
  list(
    "Baseline (21,21,21)" = ans_80_baseline,
    "Strategy 1 (21,14,14)" =  ans_80_strategy1,
    "Strategy 2 (21,7,7)" = ans_80_strategy2,
    "Strategy 3 (21,0,0)" = ans_80_strategy3
  )
)
```

### Scenario: 90% vaccination, tiered quarantine

Finally, we simulate with 90% vaccination coverage and compare the same tiered quarantine strategies:

```{r}
#| label: vaccination-90-tiered
ans_90_baseline  <- simulator(duration = c(21L, 21L, 21L), vaccinated = 0.9)
ans_90_strategy1 <- simulator(duration = c(21L, 14L, 14L), vaccinated = 0.9)
ans_90_strategy2 <- simulator(duration = c(21L, 7L, 7L), vaccinated = 0.9)
ans_90_strategy3 <- simulator(duration = c(21L, 0L, 0L), vaccinated = 0.9)

# Tabulating the results
tabulator(
  list(
    "Baseline (21,21,21)"   = ans_90_baseline,
    "Strategy 1 (21,14,14)" = ans_90_strategy1,
    "Strategy 2 (21,7,7)"   = ans_90_strategy2,
    "Strategy 3 (21,0,0)"   = ans_90_strategy3
  )
)
```

### Scenario: Lower quarantine duration (14 days)

For this scenario, we simulate with 80% vaccination coverage and a maximum quarantine duration of 14 days, comparing the same tiered quarantine strategies:

```{r}
#| label: vaccination-80-lower-duration
ans_80_21_baseline  <- simulator(duration = c(21L, 21L, 21L), vaccinated = 0.8)
ans_80_14_strategy1 <- simulator(duration = c(14L, 14L, 14L), vaccinated = 0.8)
ans_80_14_strategy2 <- simulator(duration = c(14L, 10L, 10L), vaccinated = 0.8)
ans_80_14_strategy3 <- simulator(duration = c(14L, 7L, 7L), vaccinated = 0.8)
ans_80_14_strategy4 <- simulator(duration = c(14L, 0L, 0L), vaccinated = 0.8)

# Tabulating the results
tabulator(
  list(
    "Baseline (21,21,21)"   = ans_80_21_baseline,
    "Strategy 1 (14,14,14)" =  ans_80_14_strategy1,
    "Strategy 2 (14,10,10)" = ans_80_14_strategy2,
    "Strategy 3 (14,7,7)"   = ans_80_14_strategy3,
    "Strategy 4 (14,0,0)"   = ans_80_14_strategy4
  )
)
```


## Overall comparison

Combining some of the results from different scenarios for a final comparison of the tiered quarantine strategies with 80% vaccination coverage:

```{r}
#| label: overall-comparison
tabulator(
  list(
    "Baseline (21,21,21)"   = ans_80_21_baseline,
    "Strategy 1 (14,14,14)" = ans_80_14_strategy1,
    "Strategy 2 (14,10,10)" = ans_80_14_strategy2,
    "Strategy 3 (14,7,7)"   = ans_80_14_strategy3
  )
)
```

The probability-based metrics provide a clearer picture of outbreak risk across different strategies. By examining the probability of reaching specific outbreak thresholds (10, 20, and 50 infected individuals), we can better assess the practical implications of each quarantine strategy. The distribution of total infected individuals can be further explored through density plots:

```{r}
#| label: overall-comparison-plot
# We can group the four into a single plot (histogram)
# for visual comparison
combined_results <- rbind(
  data.table(Scenario = "Baseline (21,21,21)", ans_80_21_baseline),
  data.table(Scenario = "Strategy 1 (14,14,14)", ans_80_14_strategy1),
  data.table(Scenario = "Strategy 2 (14,10,10)", ans_80_14_strategy2),
  data.table(Scenario = "Strategy 3 (14,7,7)", ans_80_14_strategy3)
)

combined_results[total_infected < 50] |>
  ggplot(aes(x = total_infected, color = Scenario)) +
    geom_density(linewidth=1.5) +
    labs(
      title = "Distribution of Total Infected Individuals - Final Comparison",
      x = "Total Infected Individuals",
      y = "Frequency"
    ) +
    theme_minimal()
```

Looking at the density plot, we can see that the ordering of the distribution in outbreak sizes is consistent with the expected impact of the different quarantine strategies.

# Discussion

The model presented here is a simplification of reality trying to explore what a change in the quarantine strategy could mean for measles outbreaks in school settings. The results suggest that quarantine duration may be reduced without significantly increasing outbreak sizes in the context of a relatively high vaccination coverage (80% or higher). However, it is important to note that the model assumes perfect compliance with quarantine measures and does not account for community transmission outside the school setting.

The mid-risk quarantine strategy--which applies to agents who are not in the same classroom but were in direct contact with an infected individual--shows a marginal effect on outbreak sizes. Nonetheless, this finding is contingent on the assumption that interactions happen randomly based on class assignments, so, in real-world scenarios where social networks and friendships play a significant role, the impact of mid-risk quarantine could be more pronounced. Because of this, it would be prudent to include mid-risk quarantine in the same category as high-risk quarantine until more detailed models are developed.

# Version

This analysis was performed using `epiworldR` version `r as.character(packageVersion("epiworldR"))`, with R version `r R.version.string`. You can get the latest version of `epiworldR` from GitHub at <https://github.com/UofUEpiBio/epiworldR>.