---
format: gfm
title: Measles Tiered Quarantine Simulations
author: George G. Vega Yon, Ph.D.
---

```{r}
#| label: setup
#| echo: false
knitr::opts_chunk$set(echo = FALSE)
```

```{=markdown}
[![](https://github.com/EpiForeSITE/software/raw/e82ed88f75e0fe5c0a1a3b38c2b94509f122019c/docs/assets/foresite-software-badge.svg)](https://github.com/EpiForeSITE/software)

> [!CAUTION]
> This project is a work in progress. Use it at your own risk. **This model simulates a single school, so community transmission is not included**.

> [!IMPORTANT]
> The model makes several assumptions that may not hold in real-world scenarios. One important assumption is that interactions between agents are based solely on class assignments, and not based on friendship networks or other social structures. This assumption reflects strongly in the effect of mid-risk quarantine (see below for more details).
```

## Description of the model

We are using the `ModelMeaslesMixingRiskQuarantine()` model from the `epiworldR` package (version 0.12.0.0 or later). This model uses a mixing matrix to represent how agents' interactions are distributed. In this case, groups represent classrooms in a school, so kids will have more interactions with kids in their own class than with kids from other classes.

The model contains the full disease progression for measles, including the incubation period, prodromal phase, and rash phase. The model assumes that kids in the rash phase are isolated from their peers. Nonetheless, the model is calibrated to reflect the measles $\mathcal{R}_0$ of 15, so kids are infectious during the prodromal period. Agents can also become hospitalized.

The quarantine process is triggered when an agent is detected during the rash period. This could occur for more than one agent in the same step. Quarantine process only applies to unvaccinated agents, and the duration of quarantine depends on the risk level of the agent, as defined below.

## Setup

This document illustrates an experiment using the `epiworldR` package to simulate measles transmission under a tiered quarantine strategy. In the tier quarantine system, agents have different quarantine durations as a function of their risk level, which are defined as follows:

- **High risk**: Agents in the same classroom as an infected individual.
- **Medium risk**: Agents who were not in the same classroom as the infected individual, but were in direct contact with them.
- **Low risk**: Agents who were not in the same classroom or direct contact with the infected individual.

Quarantine only applies to unvaccinated agents. The simulation settings are as follows:

- Individual school with 600 students distributed across 20 classes (30 students per class).
- The contact rate is given by our previous estimates for within-class and between-class interactions: 83% of contacts occur within the same class, while 17% occur between different classes.
- The basic reproduction number (R0) is set to 15, reflecting the high transmissibility of measles.
- Contact tracing is assumed to be 100% effective, as agents' willingness to isolate and quarantine.

## Results

You can see the individual results of each scenario under the folder [`scenarios/`](./scenarios/):

```{r}
#| echo: false
#| results: asis
fns <- list.files("scenarios", pattern = "[.]md", full.names = TRUE)

prop <- gsub(".+vax-([0-9.]+)[.]md", "\\1", fns)

sprintf(
  "- [Scenario with %s%% vaccination](%s)", as.numeric(prop) * 100,
  fns
  ) |>
  paste(collapse = "\n") |>
  cat()
cat("\n")
```

The overall results can be seen here:

```{r}
#| label: results
#| echo: false
#| warning: false
#| message: false
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)

fns <- list.files("scenarios", pattern = "[.]rds", full.names = TRUE)
results <- lapply(fns, \(fn) {
  x <- readRDS(fn)
  x$prop_vax <- gsub(".+scenario-([0-9.]+)[.]rds", "\\1", fn) |> as.numeric()
  x$prop_vax <- x$prop_vax / 100
  x
  }) |> rbindlist()


```

### Probability of Outbreaks by Size

The following figure shows the probability of observing outbreaks of different sizes (≥10, ≥50, and ≥100 cases) across different vaccination rates and intervention strategies.

```{r}
#| label: outbreak-probabilities
#| echo: false
#| warning: false
#| message: false
#| fig.width: 12
#| fig.height: 8
#| fig.cap: "Probability of observing outbreaks of different sizes by vaccination rate and intervention strategy"

# Calculate outbreak probabilities
outbreak_probs <- results %>%
  group_by(prop_vax, strategy) %>%
  summarise(
    prob_10_plus = mean(total_infected >= 10, na.rm = TRUE),
    prob_50_plus = mean(total_infected >= 50, na.rm = TRUE),
    prob_100_plus = mean(total_infected >= 100, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = starts_with("prob_"),
    names_to = "outbreak_size",
    values_to = "probability",
    names_prefix = "prob_"
  ) %>%
  mutate(
    outbreak_size = factor(
      outbreak_size,
      levels = c("10_plus", "50_plus", "100_plus"),
      labels = c("≥10 cases", "≥50 cases", "≥100 cases")
    ),
    vaccination_label = paste0(prop_vax * 100, "% vaccinated")
  )

# Create the visualization
ggplot(outbreak_probs, aes(x = outbreak_size, y = probability, fill = strategy)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, alpha = 0.9) +
  facet_wrap(~ vaccination_label, ncol = 2) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_fill_brewer(palette = "Set2", name = "Strategy") +
  labs(
    title = "Probability of Outbreaks by Size",
    subtitle = "Comparison across vaccination rates and intervention strategies",
    x = "Outbreak Size Threshold",
    y = "Probability"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "gray90", color = NA),
    strip.text = element_text(face = "bold", size = 11),
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "gray30")
  )
```


# Version

This analysis was performed using `epiworldR` version `r as.character(packageVersion("epiworldR"))`, with R version `r R.version.string`, and the `measles` R package version `r as.character(packageVersion("measles"))`. You can get the latest version of `epiworldR` and the `measles` R package from GitHub at <https://github.com/UofUEpiBio/epiworldR> and <https://github.com/UofUEpiBio/measles>.
